\sick, the spectroscopic inference crank

### 2014-06-09 docs/manuscript.tex
edit

We seek to maximise $\mathcal{P}$ and calculate the posterior probability distributions for $\bm{\theta}$. Continuing the example of inferring photospheric properties of the Sun, we now have a total of 16 free parameters (including $\bm{\psi}$), requiring us to interpolate between $\sim{}X\times10^{9}$ pixels (only the redder channel) in 4 dimensions. While the description might appear daunting, the problem is tractable, numerically efficient, and easy to configure: the contents of the model filename provided to \sick{} are shown in Figure \ref{fig:solar-model-file}. By default \sick{} numerically solves this problem in three sequential steps: scattering, optimisation, and Monte-Carlo Markov Chain (MCMC) sampling. These approaches are described in the following sections.

@@

In the first step we calculate the probability $\mathcal{P}$ for $N_{sample}$ randomly drawn points from all over the parameter space $\bm{\psi}$. Since \sick{} allows for arbitrarily large parameter spaces in $N_{D}$ dimensions, initially sampling the parameter space $N_{sample}$ times provides a coarse representation of probability so that we do not become `stuck' in a local minima.\footnote{If we did not numerically optimise the probability function and only sampled using MCMC with reasonable priors, becoming stuck in local minima would be less of a problem. However MCMC sampling is much more expensive than numerical optimisation.} If the probability distribution function were smoothly distributed across all $\bm{\psi}$, or the optimisation step (see below) was sufficiently robust, then in principle \textit{any} single point would be an adequate starting guess. However there is no way of knowing \textit{a priori} how smooth the probability distribution function is for a given problem.

@@

Since these optimisation algorithms are minimisation techniques and we seek to maximise the log-probability $\ln\mathcal{P}$, we numerically optimise the parameters $\bm{\theta}$ by minimising the negative log-probability $-\mathcal{P}$. After performing the initial scattering phase with $N_{samples} = 100$, the most probable sampled point is $\bm{\psi_{scat}} = [A,B,C,D]$ with $\mathcal{P} = Z$. After optimisation, the most probable values found were X.

@@

>>>>

We seek to maximise $\mathcal{P}$ and calculate the posterior probability distributions for $\bm{\theta}$. Continuing the example of inferring photospheric properties of the Sun, we now have a total of 16 free parameters (including $\bm{\psi}$), requiring us to interpolate between $\sim{}X\times10^{9}$ pixels (only the redder channel) in 4 dimensions. While the description might appear daunting, the problem is tractable, numerically efficient, and easy to configure: the contents of the model filename provided to \sick{} are shown in Figure \ref{fig:solar-model-file}. By default \sick{} numerically solves this problem in three sequential steps, which are described in the following sections: scattering, optimisation, and Monte-Carlo Markov Chain (MCMC) sampling.

@@

In the first step we calculate the log-probability $\log{(\mathcal{P})}$ for $N_{sample}$ randomly drawn points from all over the parameter space $\bm{\psi}$. Since \sick{} allows for arbitrarily large parameter spaces in $N_{D}$ dimensions, initially sampling the parameter space $N_{sample}$ times provides a coarse representation of probability so that we do not become `stuck' in a local minima.\footnote{If we did not numerically optimise the probability function and only sampled using MCMC with reasonable priors, becoming stuck in local minima would be less of a problem. However MCMC sampling is much more expensive than numerical optimisation.} If the probability distribution function were smoothly distributed across all $\bm{\psi}$, or the optimisation step (see below) was sufficiently robust, then in principle \textit{any} single point would be an adequate starting guess. However there is no way of knowing \textit{a priori} how smooth the probability distribution function is for a given problem.

@@

Since these optimisation algorithms are minimisation techniques and we seek to maximise the log-probability $\log{\left(\mathcal{P}\right)}$, we numerically optimise the parameters $\bm{\theta}$ by minimising the negative log-probability $-\log{\left(\mathcal{P}\right)}$. After performing the initial scattering phase with $N_{samples} = 100$, the most probable sampled point is $\bm{\psi_{scat}} = [A,B,C,D]$ with $\log{\left(\mathcal{P}\right)} = Z$. After optimisation, the most probable values found were X.

@@

\subsubsection{Initial Point}
\label{sec:initial-point}
If the optimisation steps is not employed, initial points for the MCMC chains are drawn in the same way as described in the scattering process. When the optimisation step is used, the initial points are taken from a small multi-dimensional ball around the optimised parameters $\bm{\theta_{opt}}$.

% Width of the ball

\begin{eqnarray}
T_{\rm eff} = 5590_{-125}^{+151}\,\mbox{K},\,\,\,\,\, \log{g} = 4.44_{-0.23}^{+0.28},\,\,\,\,\, [\mbox{Fe/H}] = -0.48_{-0.16}^{+0.21},\,\,\,\,\, [\alpha/\mbox{Fe}] = 0.02_{-0.14}^{+0.13} \\
\end{eqnarray}

% Continuum
The standard deviation of each dimension can be specified in the \sick{} model filename.
